Inequality in consumption

### 2017-02-27 inequality/consumption-inequality.Rmd
fixed error

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

@@

@@

```{r inequality_overall, cache=TRUE}
# put in both metrics of consumption
# run it through summarize_inequality
# then group by region or by drug also
usage %>%
  group_by(bene_id) %>%
  summarize(n_claims=sum(n_claims),
            n_days=sum(n_days)) %>%
  gather('consumption_metric', 'amount', n_claims, n_days) %>%
  group_by(consumption_metric) %>%
  summarize_inequality('amount') %>%
  kable
```

```{r inequality_hist, cache=TRUE}
usage %>%
  group_by(bene_id) %>%
  summarize(n_claims=sum(n_claims)) %>%
  ggplot(aes(x=n_claims)) +
  geom_histogram() +
  scale_y_log10() +
  xlab('number of claims') + ylab('number of beneficiaries') +
  theme_minimal()
```

@@

```{r inequality_by_region, cache=TRUE}
usage %>% 
  group_by(region, bene_id) %>%
  summarize(n_claims=sum(n_claims),
            n_days=sum(n_days)) %>%
  gather('consumption_metric', 'amount', n_claims, n_days) %>%
  group_by(region, consumption_metric) %>%
  summarize_inequality('amount') %>%
  kable
```

@@

```{r inequality_by_drug, cache=TRUE}
n_bene = usage$bene_id %>% unique %>% length
usage %>%
  filter(!is.na(antibiotic)) %>%
  group_by(antibiotic, bene_id) %>%
  summarize(n_claims=sum(n_claims),
            n_days=sum(n_days)) %>%
  gather('consumption_metric', 'amount', n_claims, n_days) %>%
  group_by(antibiotic, consumption_metric) %>%
  summarize_inequality('amount', n_bene) %>%
  filter(antibiotic %in% top_antibiotics) %>%
  kable
```

@@

```{r chars_by_consumption, cache=TRUE}
ntile_zero = function(x, n) {
  out = x
  out[x > 0] = ntile(out[x > 0], n)
  out
}

>>>>

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=4, fig.path='fig/',
                      echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE)
```

@@

summarize_claims = function(x) {
  summarize(x, n_claims=sum(n_claims), n_days=sum(days_supply)) %>%
    mutate(n_days=if_else(is.na(n_days), 0L, n_days))
}

@@

```{r inequality_overall}
# put in both metrics of consumption
# run it through summarize_inequality
# then group by region or by drug also
usage %>%
  group_by(bene_id) %>%
  summarize_claims() %>%
  gather('consumption_metric', 'amount', n_claims, n_days) %>%
  group_by(consumption_metric) %>%
  summarize_inequality('amount') %>%
  kable
```

```{r inequality_hist}
usage %>%
  group_by(bene_id) %>%
  summarize(n_claims=sum(n_claims)) %>%
  ggplot(aes(x=n_claims)) +
  geom_histogram() +
  scale_y_log10() +
  xlab('number of claims') + ylab('number of beneficiaries') +
  theme_minimal()
```

@@

```{r inequality_by_region}
usage %>%
  group_by(region, bene_id) %>%
  summarize_claims() %>%
  gather('consumption_metric', 'amount', n_claims, n_days) %>%
  group_by(region, consumption_metric) %>%
  summarize_inequality('amount') %>%
  kable
```

@@

```{r inequality_by_drug}
n_bene = usage$bene_id %>% unique %>% length
usage %>%
  group_by(antibiotic, bene_id) %>%
  summarize_claims() %>%
  gather('consumption_metric', 'amount', n_claims, n_days) %>%
  group_by(antibiotic, consumption_metric) %>%
  summarize_inequality('amount', n_bene) %>%
  filter(antibiotic %in% top_antibiotics) %>%
  kable
```

## Inequality by state
```{r compute_inequality_by_state}
inequality_by_state = usage %>%
  group_by(state, bene_id) %>%
  summarize_claims() %>%
  gather('consumption_metric', 'amount', n_claims, n_days) %>%
  group_by(state, consumption_metric) %>%
  summarize(n_bene=n(),
             fraction_nonzero=fraction_nonzero(amount),
             juran=juran(amount),
             juran_nonzero=juran(nonzero(amount)),
             gini=gini(amount),
             gini_nonzero=gini(nonzero(amount))) %>%
  ungroup()
```

The distribution of inequalities seems small: all states have (non-zero) Juran
indices that fall between about 0.7 and 0.75. That being said, there does seem
to be a trend by region: the fraction nonzero is highest in the South?

```{r show_inequality_by_state}
inequality_by_state %>%
  filter(consumption_metric=='n_days') %>%
  ggplot(aes(x=juran_nonzero)) + geom_histogram() +
  theme_minimal()

inequality_by_state %>%
  filter(fraction_nonzero > 0) %>%
  filter(consumption_metric=='n_days') %>%
  select(-consumption_metric) %>%
  arrange(state) %>%
  kable
```

@@

```{r chars_by_consumption}
ntile_zero = function(x, n) {
  out = x
  out[x > 0] = ntile(out[x > 0], n)
  out
}
